A Survey of Near-Term Exoskeleton Technology\\ (Draft 3.0)

### 2015-09-29 technology.tex
more updates

\input{design-goals.tex}

@@

@@

\subsubsection{What limits hydraulic performance?}

The differential pressure of the hydraulic fluid supply vs. the fluid
return limits the maximum force. Pressure times piston cross-sectional area
determines actuator force.
Many components are rated for 3000psi = 207bar (usually approximated to
210bar).
(Units: 1 Bar = $10^5$ newtons per square meter (pascals) = 14.5 psi).

@@

@@

Here is a simple actuator design:
\begin{equation}
A = 10cm^2 = 10^{-3}m^2, \; \; \; M = 0.01m
\end{equation}
Maximum torque is
\begin{equation}
P*A*M = 210*10^5N*m^{-2}*10^{-3}m^2*0.01m = 210Nm 
\end{equation}
Maximum angular velocity is
\begin{equation}
F/(A*M) = 0.2lps*(1000cm^3/l)/(10cm^2*1cm) = 20r/s
\end{equation}

@@

Can we expect a major improvement in hydraulics?
There has been incremental improvement on HPUs and valves since
the 1950s. Rapid prototyping and 3D printing are leading to
easier integration of valves, actuators, and sensors (see following
paragraph). Current high performance are flow control valves. It
is possible that lower leakage valves could be developed. It is possible
that pressure control valves would reduce the need for high bandwidth
force control, and thus high bandwidth valves. Pressure control valves
could also greatly improve exoskeleton safety if pressure control was
implemented physically rather than computationally, since physical
mechanisms are less likely to ``go unstable''.
Bypass mechnisms (a form of clutch) would allow hydraulic systems
to be backdriven without oil flow at high force, which would reduce
the power load on the HPU and the need for high performance force control.
Any valves or additional hydraulic circuitry
would have to have low internal leakage, and previous attempts at
pressure control valves have had leakage problems.
Gear shifts and variable transmissions would allow the maximum torque
and maximum no-load speed of the joint to be increased.

@@

\subsection{Reference Design: linear electric}

Survey state of the art using linear electric actuators (typically ballscrews):
RoboKnee, Roboray (Samsung)

@@

Survey state of the art using rotary electric actuators:
Schaft~\cite{shaft_foot_placement,shaft_push_recov}, Cheetah(MIT)

Maxon dc flat brushless motor EC45 and servo controller
Elmo. The servomotor emulates the kinematic profile of every
joint and sends a digital signal to the Elmo to form feedbacks.
In order to enhance the impedance capability, harmonic trans-
mission drivers are implemented (model SHD-17-100-2SH
for joints 1 and 2, model SHD-14-100-2SH for joints 3 and 4,
and CSF-32-50-2A-GR for joint 5).~\cite{IEEE07128705}

\subsection{Reference Design: linear pneumatic}

Survey state of the art using linear pneumatic actuators:

\begin{verbatim}
07139975, 07131568, 07251430.pdf
2015 IEEE International Conference on Robotics and Automation (ICRA)
Towards Balance Recovery Control for Lower Body Exoskeleton Robots
with Variable Stiffness Actuators: Spring-Loaded Flywheel Model
Corinne Doppmann, Barkan Ugurlu, Masashi Hamaya, Tatsuya Teramae,
Tomoyuki Noda, and Jun Morimoto
\end{verbatim}

\section{Sensing}

\begin{verbatim}
sensing
 actuator length/angle
 actuator velocity
 actuator force/torque 
n-line loadcells(31E-05KNO,
Sensotec) are equipped on each cylinder tube end for the
actuator force.~\cite{IEEE07222598}
 actuator sensing is pre-transmission
  joint backlash
  joint play
  structural flexibility
  actuator and transmission friction (10 Nm in Atlas)
 joint angle
three high precision(17 bit) encoders(DS-25-16, Net-
zer) are used to measure the robots joint angles(hip, knee,
ankle pitch angle).~\cite{IEEE07222598}
 joint velocity
 joint torque
 IMUs all over
 cameras all over
\end{verbatim}

{\it foot sensors, which are comprised
of four one-axis(normal force) loadcells(CBFSB-200, Bong-
shin), are adopted to detect the gait phase. The placement
of each loadcell is chosen from the analysis of GRF for
level walking : Hallux(toe), first Metatarsal(inside), fourth
Metatarsal(outside), and Heel similar to [20], as shown in
Fig. 2(a). The Gait phase detection is simply classified
into two phases by the threshold method using the sum
of the load cells: Single Stance (Left Stance(LS) or Right
Stance(RS)), Double Stance(DS). The result of gait detection
while walking is shown in Fig. 2(b).}~\cite{IEEE07222598}
20: Kong K, Tomizuka M. ”A gait monitoring system based on air pressure
sensors embedded in a shoe,” IEEE/ASME Trans Mechatron, 358 - 370.
2009.

@@

We will only address energy storage
in terms of control issues.

\subsection{Global energy storage}

\begin{verbatim}
 battery
 gasoline/jet fuel/diesel
 chemical (explosive actuation: solid fuel rocket, liquid fuel rocket)
 pressurized gas
 other?
 energy boost systems (jump, escape response in biology)
  chemical (explosive actuation: solid fuel rocket, liquid fuel rocket)
  pressurized gas
  spring
  other?
\end{verbatim}

\subsection{Joint-level/local energy storage (on joint or synergy)}

Can we use local energy storage to assist control: charging batteries,
pressurizing accumulators, using springs, ... ?

{\it incorporation of physical
compliance is of importance when it comes to dependability,
low mechanical impedance, locomotion efficiency, inherent
safety and enhanced environmental interaction capabilities}~\cite{IEEE07139975}

\begin{verbatim}
 accumulator (really an air spring)
 physical springs/damping
 supercapacitor/battery/electrical for local electric braking
\end{verbatim}

@@

Survey possible actuation systems.

@@

\subsection{Global actuation}

\begin{verbatim}
 hydraulic pump 
 gas (air) pump
\end{verbatim}

\subsection{Joint-level/synergy/local actuation}

\begin{verbatim}
 rotary electric
  motor
 linear electric
  ball screw (both motor and transmission)
 linear hydraulic
  moog thing
 rotary hydraulic
 linear pneumatic
 rotary pneumatic
 linear or rotary impulsive actuation (IC engine style actuation)
 energy boost systems (jump, escape response in biology)
  chemical (air bag style: explosive actuation: solid/liquid fuel rocket)
  pressurized gas
  spring
  other?
\end{verbatim}

@@

Energy storage and actuation for high manueverability.

@@

\begin{verbatim}
parallel combinations - weak fast actuator needs to hit joint limit and then be strong.
 (leads to chattering) or just lock in position to exert big force
series combination - strong actuator needs to clutch out of circuit to move fast
\end{verbatim}

\section{Transmissions and Drive Trains}

\subsection{How can clutches and gear shifts help?}

How can we implement a gear shifting system (the crucial mechanical
system missing in current robots)? How do we switch from manipulation
while standing, walking, running, other rapid movement (dodging), and
power movements (jumping, moving through doors and walls)?

Stance leg requires a large torque and
relatively low bandwidth during the support period. However,
the swing leg undergoes a large motion, but it only supports
its own weight during the swing period. Therefore, the swing
leg requires a relatively small torque. Thus, a dual-mode
approach is very useful for an exoskeleton robot control.
In this paper, a dual-mode control method using a hydraulic
system is proposed,~\cite{IEEE07222598}

Rise from squat, chair.
push large objects out of the way (standing full body push
or seated using legs to push)

\subsection{Global Transmission}

\begin{verbatim}
 electric circuit
 hydraulic circuit
 gas (air) circuit
\end{verbatim}

\subsection{Joint-level/synergy/local transmission}

\begin{verbatim}
 rigid
  gears
   planetary
   other
  harmonic drive
  ball screw
  cosine attachment
  Sarcos Primus Humanoid knee - to reduce moment arm variation with angle
 Bowden cable
 tendon drives (can go slack)
 hydraulic/pneumatic circuit
 variable
  belts
  cones
  fluid
  adjustable moment arm
   linear slider
   XY slider for actuation and ``joint'' on foot.
 series elastic actuation
\end{verbatim}

\section{Address what is the complexity penalty, weight penalty, 
and ``drag'' of actuator on individual joint motion.}

\section{System issues}

Things that don't fit well elsewhere.

\begin{verbatim}
 bearing play, structural flexibility: +/- 2cm COM uncertainty
 how attached to human (exoskeleton)
 air bags
\end{verbatim}

The human and exoskeleton limbs are typically connected to each
other with straps or padding. These connections typically have substantial
compliance and play, and may or may not have appropriate damping.
Typically the human and the exoskeleton do not move the same way:
pose, and linear and angular velocities and accelerations are not the same.

Heat exhaust. Heatsink cooling, Force air cooling, Water cooling.
Will housing and structure get too hot and fry operator?

@@

@@

1) We recommend the exploration of how to implement and utilize practical
gear shifts and clutches (or equivalents). For hydraulics and air, this might mean
bypass or other special valve features.
For electric motors and gears, this might mean special purpose clutch mechanisms
built into gearing or harmonic drives. For linear actuators, antagonistic actuation
combined with tendons that can go slack can be used for clutching.

2) We recommend exploring the combination of passive energy storage (springs,
accumulators, supercapacitors, ...) with active actuation. One form of this
is series elastic actuation.

3) We also recommend exploring storing energy with
a ``catch'' and ``explosive release'' for overdrive capabilities.

4) We recommend the exploration of hybrid actuation, where slow and strong actuators
are combined with fast but weak actuators either in parallel or serial.

5) We recommend the exploration of multi-joint or synergy actuation, where an actuator
might cross two or more joints.

6) We recommend the exploration of integrated actuation units using 3D printing
or other rapid prototyping techniques, such as the integrated Moog hydraulic
actuator for HyQ.

7) We recommend applying as many sensors as possible, and asking the performers
to make it easy to add more by making the sensor network available in the design.
This maximizes the probability of success
by enabling multiple control
approaches to be implemented, refined, and support each other.
In particular, we would like to see force sensing between the operator and the
exoskeleton, including at the feet, as much force sensing between the exoskeleton 
and the world as possible, but minimally full six-axis force/torque sensing at
the exoskeleton feet (where they touch the ground). We would like to see multiple MEMs
IMUs (measuring linear acceleration and angular velocity) installed across 
the exoskeleton and at least one high quality fiber optic gyro IMU.
We would like to see high quality direct velocity sensing
such as high count (100,000 counts/revolution) encoders and/or 
analog rotary or linear tachometers on actuators and joints. 
We would like to see actuator force or torque sensors
(load cells or equivalents) with the measurement on the link side (rather than
the actuator side) of any transmission. Electric current in motors and oil
pressure in hydraulic pistons can also be used for actuator force estimation,
but because these measurements are on the actuator side before the transmission
and in the case of hydraulics before the oil seals on the piston, these measurements
are greatly contaminated by friction. On the Atlas humanoid we typically saw 10Nm
joint torque estimation errors for a system that estimated actuator output using
oil pressure on each side of the piston head.
A possible objection to this is additional cost. We feel it is a false
economy to skimp on sensing. Leaving practical sensing out greatly increases
the risk of poor performance.

8) Do we recommend flywheels?

More Conclusions and Recommendations to be written.

>>>>

%\input{design-goals.tex}

@@

\section{Specifications}

We would like to find actuation that can keep up with human
speeds and provide forces or torques comparable to human capabilities:
\begin{enumerate}
\item
2-4 revolutions/second maximum joint speed.
\item
200-400Nm maximum torque at a joint.
\end{enumerate}

{\it Dennis E. Anderson, Michael L. Madigan, Maury A. Nussbaum,
“Maximum voluntary joint torque as a function of joint angle and
angular velocity: Model development and application to the lower
limb”, Journal of Biomechanics, 2007, Volume. 40, Issue 14, Pages
3105-3113} says max knee torque is around 350Nm. They only consider
max knee angular velocities around 2revolutions/second.

@@

This section considers hydraulics using linear actuators (pistons).
It is possible that rotary hydraulics, such as vane motors,
can do better, but we are not aware of non-proprietary examples
to discuss.

\subsubsection{What limits performance?}

The differential pressure of the hydraulic fluid supply vs.\ the fluid
return limits the maximum force. Pressure times piston cross-sectional area
determines actuator force.
Many components are rated for 3000psi = 207bar (usually approximated to
210bar).
(Units: 1 bar = $10^5$ newtons per square meter (pascals) = 14.5 psi).

@@

The size and weight of an HPU, and how to power it are limiting.
There have been examples of portable HPUs driven by electric and IC motors.
The Atlas humanoid
robot had an HPU driven by batteries (don't have specs).
Max pressure was around 2600psi.
We have specifications for a portable 
5kW HPU with 4gpm at 3000psi; 170V, 30A input; weighing
40lb and 15x10x10in in size.
A heat exchanger is an additional 14x28x6in in size.
Batteries to power this would be also take up space and 
add weight.

\subsubsection{A reference design}

@@

Here is a simple actuator design:
\begin{equation}
A = 10cm^2 = 10^{-3}m^2, \; \; \; M = 0.01m
\end{equation}
Maximum torque is
\begin{equation}
P*A*M = 210*10^5N*m^{-2}*10^{-3}m^2*0.01m = 210Nm 
\end{equation}
Maximum angular velocity is
\begin{equation}
F/(A*M) = 0.2lps*(1000cm^3/l)/(10cm^2*1cm) = 20r/s = 3.2rev/s
\end{equation}

@@

\subsubsection{What have we learned from implementations?}

High performance force control has been demonstrated many times
on many hydraulic robots (Sarcos, Atlas, HyQ, ...)

However, on the Atlas robot differential oil pressure is used
to estimate actuator force, as this form of force sensing
is quite mechanically robust (low maintenance).
This measurement is before the piston
seals and transmission, so it does not capture those sources
of friction. The joint torques are off by up to 10Nm at each joint.
The Sarcos robot uses load cells in the actuator to joint transmissions.

On the Atlas humanoid robot there was electromagnetic interactions between
nearby valves.

\subsubsection{Can we do better?}

Can we expect major improvements in hydraulics?
There has been incremental improvement on HPUs and valves since
the 1950s. Rapid prototyping and 3D printing are leading to
easier integration of valves, actuators, and sensors (see following
paragraph). Current high performance valves are flow control valves. It
is possible that lower leakage valves could be developed. It is possible
that pressure control valves would reduce the need for high bandwidth
force control, and thus high bandwidth valves. Pressure control valves
could also greatly improve exoskeleton safety if pressure control was
implemented physically rather than computationally, since physical
mechanisms are less likely to ``go unstable''.
Bypass mechanisms (a form of clutch) would allow hydraulic systems
to be backdriven without oil flow at high force, which would reduce
the power load on the HPU and the need for high performance force control.
Any valves or additional hydraulic circuitry
would have to have low internal leakage, and previous attempts at
pressure control valves have had leakage problems.
Gear shifts and variable transmissions would allow the maximum torque
and maximum no-load speed of the joint to be increased (also true
for all other actuation types).

@@

@@

This section considers heavily geared rotary electric motors.

\subsubsection{What limits performance?}

Electric motors are much more flexible than hydraulics in terms of voltages
and currents, as there is not a standard voltage equivalent to the 3000psi
design point in hydraulics.

Ultimately, electric motors, gears, and other drive train elements such
as clutches and variable transmissions are limited by size and weight.

One limitation on electrical current
is thermal. Motors overheat and windings melt.
A typically 
higher limit on electrical current is demagnetization of motor magnets.

\subsubsection{A reference design}

The JST lab at the University of Tokyo pushed the design of water
cooled electric motors and motor drivers. This led to the high
performance of the Schaft robot, a spinoff from that lab.

The water cooled motor driver they developed is capable of 80V and
80A continuous.~\cite{IEEE05649683}

They modified
MAXON EC-powermax30 200W brush-less motors, and added harmonic drives
to make an integrated actuator.
The maximum speed of these actuators was 3 revolutions/second.
The maximum torque using 2 drivers/motor was theoretically 400Nm, but
we have only seen 300Nm in the published data. 2 rev/s can be achieved
at peak torque (Fig~\ref{f:s1}).~\cite{IEEE05649683}

\begin{figure}[t]
\centering
\includegraphics[width=0.75\textwidth]{tech-figs/s1}
\caption{Water cooled actuator capabilities.}
\label{f:s1}
\end{figure}

Peak currents are in the range of 100-150A.
{\it
``High current load on Li-ion or Ni-MH batteries affects life
span or cause damage. They cannot be charged with high
current rate. Electrical double layer capacitors are suitable for
such instantaneous current supply because they can supply
high current without degradation for high current while
their power capacity is not large. Capacitors can be charged
with high current. Combination of an electrical double layer
capacitor and battery is adopted. The robot carries a capacitor
developed for it by Shizuki Electric Co. Inc. Specifications of
the developed capacitor are shown in Tab.II.''}~\cite{IEEE05649683}
Note that this capacitor weighs 9 kg (Fig~\ref{f:s2}).

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{tech-figs/s2}
\caption{Capacitor specifications.}
\label{f:s2}
\end{figure}

\subsubsection{What have we learned from implementations?}

This effort ``went dark'' after Google bought the spinoff company
Schaft, so there is a lot we don't know.

There are concerns about motor lifetime with this pattern of utilization.

This approach relies on transient loading of the motor, which needs
to be planned in advance to avoid not having thermal capacity
when more high performance is needed.
It is not clear how much this approach helps with 
steady state behaviors such as running.

Driving the motors this hard generates a lot of electromagnetic interference,
so they went with optical fiber for information transmission in their
robots.

We don't know whether these actuators can be made backdrivable with
high performance torque control. Pre-Schaft, the researchers were exploring
nonlinear series elastic actuation (big springs)
to help handle impacts and large force
transients~\cite{IEEE06651562}.

\subsubsection{Can we do better?}

This idea is a relatively new one, so there is probably useful unexplored
territory in the design space.

@@

We will only address energy storage
in terms of control issues. We are particularly interested in the
possibilities of joint-level or local energy storage.

Systems whose global energy storage is batteries waste
less energy driving electrical systems, but there are plenty of electrically
driven hydraulic systems on robots.
Systems whose global energy storage is liquid fuel (gasoline etc)
waste less energy driving a hydraulic system.

\subsection{Joint and synergy-level energy storage}

Can we use local energy storage to assist control: charging batteries or
supercapacitors, pressurizing accumulators, using springs, ... ?
The reference electrical design used a big capacitor to help handle
peak loads. We believe that many behaviors ``slosh'' potential and
kinetic energy back and forth.
Running exchanges vertical velocity and energy storage in spring-like
tissue.
Walking exchanges forward velocity and gravitational storage.
We also believe that energy storage can be used to start and stop
highly dynamic transient behaviors.
Jumping stilts and pogo sticks are an excellent demonstration
of the use of energy storage to enhance human
capabilities.
\url{https://www.youtube.com/watch?v=lbp41vWP4o4},
\url{https://www.youtube.com/watch?v=9ZOd7yEyhwI}

In addition, physical springs protect the exoskeleton from damage
and contribute to system dependability.
The JST lab was exploring human-scale jumping robots
with 
nonlinear series elastic actuation (big springs)
to help handle impacts and large force
transients~\cite{IEEE06651562}.

@@

@@

Here are some actuation types that have not been adequately explored
in humanoid robots and exoskeletons.

@@

Energy storage and actuation for superhuman manueverability.
\url{http://www.ric.org/app/files/public/4898/LemusVallery_GyroscopicBackpack_EMBC2014pdf.pdf}
\url{http://www.ncbi.nlm.nih.gov/pubmed/22169384}

\subsection{Internal combustion actuators}

Several labs have explored using the piston of an internal combustion
engine as a stand-alone linear or rotary actuator. In this case
the combustion event would be carefully metered and timed. Repetitive
combustion events combined with mechanical filtering could lead to smooth
output forces. This approach inherits the desirable energy storage capacity
of liquid fuels. 
\url{http://softroboticstoolkit.com/book/combustion-driven-actuators}
\url{http://www.nature.com/articles/srep04296}
\url{http://arxiv.org/abs/1402.7101}
\url{http://ntrs.nasa.gov/archive/nasa/casi.ntrs.nasa.gov/20100021938.pdf}
% https://smartech.gatech.edu/bitstream/handle/1853/19873/warta_brett_j_200712_mast.pdf
\url{http://www.msoe.edu/servlet/JiveServlet/previewBody/5589-102-1-7785/Sarah_Swanson_Final_Paper.pdf}

@@

We can consider the following combinaitons:
\begin{itemize}
\item
parallel combinations - fast but weak  actuator needs to hit joint limit and then be strong. (leads to chattering) or just lock in position to exert big force
\item
series combination - strong but slow actuator needs to clutch out of circuit to move fast
\end{itemize}

\section{How can clutches and gear shifts help?}

How can we implement a gear shifting system (the crucial mechanical
system missing in current robots)? How do we switch from manipulation
while standing, walking, running, other rapid movement (dodging), and
power movements (getting up from a squat or chair,
jumping, moving through doors and walls, pushing large objects
with full body while standing or seated)?

Examples:~\cite{IEEE07222598}

@@

{\bf Non-rigid systems:}
Light exoskeletons will have a lot of structural flexibility,
and will be poorly defined kinematically. This is not a
problem for applying forces, but it is a problem for determining
positions and velocities, such as center of mass position and velocity.

{\bf How exoskeleton attached to human:}
The human and exoskeleton limbs are typically connected to each
other with straps or padding. These connections typically have substantial
compliance and play, and may or may not have appropriate damping.
Typically the human and the exoskeleton do not move the same way:
pose, and linear and angular velocities and accelerations are not the same.

{\bf Heat exhaust:}
Will actuators, housing, and structure get too hot and fry operator?
Heatsink cooling, Force air cooling, Water cooling.
Hot material ejection.

@@

1) We recommend the exploration of how to implement and utilize practical
{\bf gear shifts and clutches} (or equivalents). For hydraulics and air, this might mean
bypass or other special valve features.
For electric motors and gears, this might mean special purpose clutch mechanisms
built into gearing or harmonic drives. For linear actuators, antagonistic actuation
combined with tendons that can go slack can be used for clutching.

2) We recommend exploring the combination of {\bf passive energy storage} (springs,
accumulators, supercapacitors, ...) with active actuation. One form of this
is series elastic actuation.

3) We also recommend exploring storing energy with
a ``catch'' and ``explosive release'' for {\bf overdrive capabilities}.

4) We recommend the exploration of {\bf hybrid actuation,} where strong but slow actuators
are combined with fast but weak actuators either in parallel or serial.

5) We recommend the exploration of {\bf multi-joint or synergy actuation}, where an actuator
might cross two or more joints.

6) We recommend the exploration of {\bf modular integrated sensing/actuation units} using 3D printing
or other rapid prototyping techniques, such as the integrated Moog hydraulic
actuator for HyQ.

7) We recommend applying {\bf as many sensors as possible,} and asking the performers
to make it easy to add more by making the sensor network available in the design.
This maximizes the probability of success
by enabling multiple control
approaches to be implemented, refined, and support each other.
In particular, we would like to see force sensing between the operator and the
exoskeleton, including at the feet, as much force sensing between the exoskeleton 
and the world as possible, but minimally full six-axis force/torque sensing at
the exoskeleton feet (where they touch the ground). We would like to see multiple MEMs
IMUs (measuring linear acceleration and angular velocity) installed across 
the exoskeleton and at least one high quality fiber optic gyro IMU.
We would like to see high quality direct velocity sensing
such as high count (100,000 counts/revolution) encoders and/or 
analog rotary or linear tachometers on actuators and joints. 
We would like to see actuator force or torque sensors
(load cells or equivalents) with the measurement on the link side (rather than
the actuator side) of any transmission. Electric current in motors and oil
pressure in hydraulic pistons can also be used for actuator force estimation,
but because these measurements are on the actuator side before the transmission
and in the case of hydraulics before the oil seals on the piston, these measurements
are greatly contaminated by friction. On the Atlas humanoid we typically saw 10Nm
joint torque estimation errors for a system that estimated actuator output using
oil pressure on each side of the piston head.
A possible objection to this is additional cost. We feel it is a false
economy to skimp on sensing. Leaving practical sensing out greatly increases
the risk of poor performance.
